---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# trendfilter

<!-- badges: start -->
[![R-CMD-check](https://github.com/glmgen/trendfilter/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/glmgen/trendfilter/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of trendfilter is to **solve** nonparametric regression.

## Installation

You can install the development version of trendfilter from 
[GitHub](https://github.com/) with:

```{r}
# install.packages("remotes")
remotes::install_github("zhoumo2716/trendfilter")
#pak::pak("zhoumo2716/trendfilter")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r}
library(trendfilter)
library(ggplot2)
x <- 1:100 / 101 * 2 * pi
y <- sin(x) + .2 * rnorm(100)
out <- trendfilter(y, x, nlambda = 15)
plot(out) +
  geom_point(data = data.frame(x = x, y = y), aes(x, y), color = "black")
```

## Boundary Condition Option

`boundary_condition = TRUE` enables boundary correction using Newton polynomials and divided differences.
`left_boundary_m` and `right_boundary_m` determine how the boundary conditions are applied (order of polynomial). By default, `left_boundary_m` = `right_boundary_m` = `round(k/2)`, which simplifies to (k+1)/2 when k is odd (natural spline at the boundary).

```{r}

# Standard trend filtering (without boundary conditions)
out1 <- trendfilter(y, x, nlambda = 3, lambda_min = 0.1)
plot(out1) +
  ggtitle("Standard Trend Filtering") +
  geom_point(data = data.frame(x = x, y = y), aes(x, y), color = "black")

# Trend filtering with natural spline boundary conditions
out2 <- trendfilter(y, x, nlambda = 3, lambda_min = 0.1, boundary_condition = TRUE)
plot(out2) +
  ggtitle("Natural Spline Boundary") +
  geom_point(data = data.frame(x = x, y = y), aes(x, y), color = "black")

# Custom boundary conditions
out3 <- trendfilter(y, x, nlambda = 3, lambda_min = 0.1, boundary_condition = TRUE, left_boundary_m = 2, right_boundary_m = 1)
plot(out3) +
  ggtitle("Custom Boundaries (L=2, R=1)") +
  geom_point(data = data.frame(x = x, y = y), aes(x, y), color = "black")




```
